PAMELA: an annotation-based Java Modeling Framework

### 2021-02-17 introduction.tex
introduction after Antidote

Model-Driven Software Development focuses on managing abstract models representing the conceptual level of the solution space. These models are generally represented as various artifacts, corresponding to different languages and using different representations. But models are rarely executable and/or lack the required level of expressiveness or performance to be exploited directly in production. Thus, the solution they offer is often implemented by using code generators that produce a representation of the solution in a target programming language or framework\cite{stahl2006model}. The underlying semantics of the code to be executed is generally encoded in those code generators and can be inlined or implicitly defined by the code generation process, or may be sometimes explicit to the code generation. Figure \ref{fig:ClassicalVision} shows the classical vision for Model-Driven Engineering.

@@

Round-trip mechanisms are commonly used to overcome those difficulties, but they are difficult to use and to maintain. Indeed, the model/code co-evolution problem remains an open subject in the software engineering research community and, consequently, software development projects often abandon the idea of maintaining the synchronization between model and source code during development process. In that context, the model is developed at the early stages of development process, used mainly to prototype software application. Then, it may be manually reviewed back at the end of development process, for documentation purposes.

\begin{figure}
    \centering
    \includegraphics[width=1.0 \columnwidth]{ClassicalVision.pdf}
    \caption{Classical vision for Model-Driven Engineering}
    \label{fig:ClassicalVision}
\end{figure}

>>>>

Model-Driven Software Development focuses on managing abstract models representing the conceptual level of the solution space. These models are generally represented as various artifacts, corresponding to different languages and using different representations. But models are rarely executable and/or lack the required level of expressiveness or performance to be exploited directly in production. Thus, the solution they offer is often implemented by using code generators that produce a representation of the solution in a target programming language or framework\cite{stahl2006model}. The underlying semantics of the code to be executed is generally encoded in those code generators and can be inlined or implicitly defined by the code generation process, or may sometimes be explicit to the code generation. Figure \ref{fig:ClassicalVision} shows the classical vision for Model-Driven Engineering.

@@

Round-trip mechanisms are commonly used to overcome those difficulties, but they are difficult to use and to maintain. Indeed, the model/code co-evolution problem remains an open subject in the software engineering research community and, consequently, software development projects often abandon the idea of maintaining the synchronization between model and source code during development process. In that context, the model is developed at the early stages of development process, used mainly to prototype software applications. Then, it may be manually reviewed back at the end of development process, for documentation purposes.

\begin{figure}
    \centering
    \includegraphics[width=1.0\columnwidth]{ClassicalVision.pdf}
    \caption{Classical vision for Model-Driven Engineering}
    \label{fig:ClassicalVision}
\end{figure}
